---
title: "p8105_hw1_ys3766"
author: "Yifan Shi"
date: "2024-09-21"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggplot2)
```

The purpose of this file is to present my work for Homework 1


## Problem 1

Load the penguins dataset
```{r}
data("penguins", package = "palmerpenguins")
```

Describe the dataset
```{r}
str(penguins)
ncol(penguins)
nrow(penguins)
summary(penguins)
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

The 'penguins' is a dataset from the 'palmerpenguins' package. It includes `r nrow(penguins)` observations and `r ncol(penguins)` variables. It contains variable "species", "island", "bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g", "sex" and "year". 

The dataset contains data for 344 penguins of 3 species, namely Adelie, Chinstrap and Gentoo. They are from 3 different islands, Biscoe, Dream and Torgersen. The mean flipper length of the penguins is 200.9mm.


Scatterplot of flipper_length_mm vs bill_length_mm

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
	geom_point()+
  labs(title = "Scatterplot of Penguin Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species")+
  theme_minimal()
ggsave("scatterplot.png")
```

## Problem 2
Create a data frame comprised of:
  * a random sample of size 10 from a standard Normal distribution
  * a logical vector indicating whether elements of the sample are greater than 0
  * a character vector of length 10
  * a factor vector of length 10, with 3 different factor “levels”
```{r}
df_hw1 = tibble(
  norm = rnorm(10),
  vec_logical = norm > 0,
  vec_character = letters[1:10],
  vec_factor = factor(rep(c("one", "two", "three"), length.out = 10))
)
```

